
\chapter{Conclusion}
\label{ch:conclusion}

This master thesis addressed a research gap in applying training-free and schema-agnostic \gls{kgqa} approaches to the task of scholarly literature search. Our research aimed to overcome the limitations associated with existing approaches, specifically their reliance on fixed schemas and training data (\hyperref[sec:problem_statements]{\textbf{P1}}), and the lack of a standardized taxonomy for assessing capabilities of \gls{kgqa} systems (\hyperref[sec:problem_statements]{\textbf{P2}}).

Consequently, the master thesis introduced HubLink (\hyperref[enum:c1]{\textbf{C1}}), a novel approach that represents a significant step toward schema-agnostic and training-free \gls{kgqa} retrieval. By conceptually decomposing the graph into structures termed \emph{hubs}, which aggregate the knowledge from scholarly publications, HubLink enables a modular and source-aware retrieval process. Specifically, the ability of the approach to transparently trace the origin of information during inference provides a key requirement for scholarly literature retrieval. Furthermore, the embedding-based methodology employed by HubLink leverages the semantic connections captured by \glspl{llm} without necessitating explicit training data or adhering to fixed graph schemas. This independence makes the approach inherently adaptable to dynamic and evolving graph structures. 

Our evaluation results demonstrate that HubLink substantially improves the relevance and accuracy of retrieved contexts compared to state-of-the-art baseline \gls{kgqa} approaches, many of which struggled in the scholarly domain application. The evaluation further highlights that HubLink can be successfully applied to six specific use cases for the scholarly literature task, showing the potential of the approach to improve the efficiency of scientific communication. Although the evaluation demonstrated the clear advantages of HubLink in retrieving relevant information, it also identified specific areas for future enhancement. In particular, the results suggest that there are opportunities to refine the precision of the retrieved context and their ranking by relevance. Furthermore, the process of generating coherent and precise responses based on the retrieved context warrants further investigation and development. Future work should therefore focus on improving these aspects to fully realize the potential of the approach.

Complementing the HubLink approach, this thesis proposed a new taxonomy for \gls{kgqa} specifically tailored to the literature search task (\hyperref[enum:c2]{\textbf{C2}}). Developed through a systematic and operationalized construction process (\hyperref[enum:c2]{\textbf{C2.2}}), this taxonomy provides a structured framework to classify questions based on characteristics relevant to retrieval performance and scholarly information needs. The validation process, which includes the application to research questions and comparison with existing dataset-specific classifications, confirms that the taxonomy addresses a notable gap by offering a more structured and fine-grained framework for understanding and assessing \gls{kgqa} system capabilities in the scholarly domain. The usefulness of this taxonomy is further demonstrated through its successful application in the construction of new \gls{kgqa} datasets (\hyperref[enum:c3]{\textbf{C3}}) for the \gls{orkg}. These datasets were used to conduct the comprehensive evaluation of HubLink and the baseline methods across diverse question types. 

To conclude, current scholarly practices embed scholarly findings within unstructured documents, rendering data extraction and targeted search cumbersome. In contrast, \glspl{rkg} store scholarly findings in a structured network. The application of HubLink to retrieve relevant data from \glspl{rkg} in a \gls{kgqa} setting enables direct and precise access to scholarly information, reducing manual effort. This paradigm significantly reduces the effort required by researchers to locate relevant scholarly findings. Consequently, this work makes a significant contribution to the advancement of research in scholarly \gls{kgqa}, offering a promising path towards more efficient and effective scientific discovery and communication in the future.

\input{chapters/11_conclusions/sections/research_questions_revisited}
\input{chapters/11_conclusions/sections/future_work}
% \input{chapters/10_conclusions/sections/conclusion}


% Die ERstellung der QA Datasets war schwieriger als erwartet und hat dardurch deutlich mehr Zeit in Anspruch genommen als geplant. Für die Zukunft wäre es sinnvoll, die LLM gestütze Generierung von KGQA Datasets genauer zu untersuchen. Die Herausforderungen und Requirements sollten empirisch bestimmt werden.

